---
title  : "Risk of COVID-19 related hospitalisation and death post primary vaccination schedule and booster dose in Wales"
date   : "Date compiled: `r format(Sys.time(), '%a %d %B %Y')`"
author : "[Stuart Bedston](mailto:stuart.bedston@swansea.ac.uk)"
---

```{r setup, include=FALSE}

source("r_clear_and_load.r")

options(
    knitr.kable.NA = ''
)

knitr::opts_chunk$set(
    fig.width  = 6,
    fig.height = 5,
    fig.align = "center",
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)

sep21 <- ymd("2021-09-01")

too_small <- function(x) {ifelse(x < 1/10^9, -Inf, x)}
too_big   <- function(x) {ifelse(x > 10^9, Inf, x)}

tidy_coef <- . %>%
    mutate(
        coef = str_glue("{est} ({low}, {high})",
            est  = format(round(estimate,  2), nsmall = 2),
            low  = format(round(too_small(conf_low),  2), nsmall = 2),
            high = format(round(too_big(conf_high), 2), nsmall = 2)
        )
    ) %>%
    select(xvar, xlbl, coef)

```

## 0. Message to SAIL reviewers

Please start with “99_report.html” before checking the CSVs, it gives context to the analysis, explains the data sources used, the suppression applied to each bit of output, and lists how the CSVs are relevant.

In general, for Figures, all counts with values between 1 and 9 have been replaced with 10.
Whilst for Tables they have been omitted, with all other counts rounded to nearest 10.
For model parameter estimates, no suppression has been applied.

## 1. Research aims

Identify demogrpahic and health charcteristics assoiated with greater risk
of COVID-19 related hospitalisation and death for those in the general
populaton who have revcieved a second or booster COVID-19 vaccination

## 2. Study and analysis design

Prospective observational study on a national e-cohort of those alive and
living in Wales from December 2020 onwards who have at least had their
second dose under the COVID-19 primary vaccination schedule.

### Data sources

- **C19_COHORT20**:
    - **WDSD**: Welsh Demographic Service Dataset
    - **WLGP**: Welsh Longitudinal General Practice data
- **C19_COHORT_MORTALITY**:
    - **ADDE**: Annual District Death Extract
    - **CDDS**: COVID-19 Consolidated Deaths Data Set
    - **WDSD**: Welsh Demographic Service Dataset
- **(RRDA_)CVVD**: COVID-19 vaccination records from the Welsh Immunisation System
- **(RRDA_)HWRA**: Health Workers list in NHS Wales organisations
- **PATD**: COVID-19 PCR test results
- **PEDW**: Patient Episode Dataset for Wales
- **WDSD**: Welsh Demographic Service Dataset
- **WLGP**: Welsh Longitudinal General Practice data
- **CENW**: ONS census for Wales

### Table 2.1: Cohort selection

- Counts rounded to nearest 10, percentages to 1 decimal place

```{r}

qread("results/t_cohort_selection.qs") %>%
mutate_at(c("n", "n_diff"), round, digits = -1) %>%
mutate(p_diff = round(p_diff, 1)) %>%
kable(
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)

```

## 3. Cohort from Dec 2020

```{r}

qload("results/cohort_freq_plots.qsm")
qload("results/cuminc_uptake.qsm")

```

### Figure 3.1: Weekly variant proportion

- No suppression needed as the plots are based on data from outside of SAIL

```{r, fig.height = 3}

qload("results/variant.qsm")

p_variant_week_freq

p_variant_week_prop

```

### Figure 3.2: Weekly infection counts, overlaid with background variant and changepoints in rate

- Counts between 1 and 9 replaced with 10

```{r}
qread("results/p_bg.qs")
```

### Figure 3.3: Vaccination counts by week

- Counts are thousands per week
- Counts between 1 and 9 replaced with 10

```{r}
p_vacc_week
```

### Figure 3.4: Number of weeks between vaccinations

- Counts between 1 and 9 replaced with 10

```{r}
p_vacc_diff
```

### Figure 3.5: Number of positive PCR tests and hospitalisation/deaths by week

- Counts between 1 and 9 replaced with 10

```{r}
p_infect_hosp_death
```

### Figure 3.6: Method of hospital admission for COVID-19 admissions

- Counts between 1 and 9 replaced with 10

```{r}
qread("results/p_hosp_admis.qs")
```

### Figure 3.7: COVID-19 status for emergency COVID-19 hospital admissions

- Counts between 1 and 9 replaced with 10

```{r}
qread("results/p_hosp_emg_c19.qs")
```

### Figure 3.8: Weeks between vaccination and positive PCR

- Counts between 1 and 9 replaced with 10

```{r}
p_infect2dose
```

### Figure 3.9: Cumulative incidence of vaccination status

```{r, fig.height = 10}
p_vacc_state
```

# 4. Analysis sample selection

### Table 4.1: Analysis sample selection

- Counts rounded to nearest 10
- Diff columns calculated after rounding

```{r}
t_analysis_sample_selection <- qread("results/t_analysis_sample_selection.qs")

t_dose2_sample_size <-
    qread(s_drive("d_analysis_dose2.qs")) %>%
    summarise(
        total_n = n_distinct(alf_e),
        outcome_n = sum(event_flg)
    )

t_doseb_sample_size <-
    qread(s_drive("d_analysis_doseb.qs")) %>%
    summarise(
        total_n = n_distinct(alf_e),
        outcome_n = sum(event_flg)
    )

# print
t_analysis_sample_selection %>%
mutate(across(
    .cols = matches("_n"),
    .fns = ~ if_else(
        condition = 1 <= abs(.x) & abs(.x) <= 9,
        true = .x / abs(.x) * 10,
        false = round(.x, -1),
        missing = NA_real_
    )
)) %>%
mutate(
    total_n_diff = total_n - lag(total_n),
    outcome_n_diff = outcome_n - lag(outcome_n)
) %>%
kable(
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)
```

# 5. Analysis of break through for dose 2 and booster from 13 Sept 2021

### Table 5.1: Descriptive summaries for Dose 2 in total

- N, person-years, event counts, and rates are omitted if N or event count is between 1 and 9
- N and event counts rounded to nearest 10
- Person-years and rate per 1,000 person years rounded to 1dp
- The following CSV is the raw version of Tables 5.1a, 5.1b, 5.1c combined: `t_desc_dose2b_rate.csv`

```{r}
t_desc_dose2b_rate <-
    qread("results/t_desc_dose2b_rate.qs") %>%
    mutate(
        xvar = fct_inorder(xvar),
        xlbl = fct_inorder(xlbl)
    )

# add totals to table
t_desc_dose2b_total <-
    t_desc_dose2b_rate %>%
    group_by(xvar, xlbl) %>%
    summarise(
        n = sum(n),
        pyears = sum(pyears),
        event = sum(event)
    ) %>%
    ungroup() %>%
    mutate(.after = xlbl, vacc_dose2_name = "Total") %>%
    mutate(rate = event / pyears * 1000)

t_desc_dose2b_rate <-
    bind_rows(
        t_desc_dose2b_total,
        t_desc_dose2b_rate,
    )

# suppression
t_desc_dose2b_rate <-
    t_desc_dose2b_rate %>%
    mutate(
        n      = ifelse(1 <= n & n <= 9, NA, n),
        n      = round(n, -1),
        event  = ifelse(1 <= event & event <= 9, NA, event),
        event  = round(event, -1),
        pyears = ifelse(is.na(n), NA, pyears),
        pyears = round(pyears, 1),
        rate   = ifelse(is.na(event) | is.na(pyears), NA, rate),
        rate   = round(rate, 1)
    )

# print Total
t_desc_dose2b_rate %>%
filter(vacc_dose2_name == "Total") %>%
filter(xvar != "health_board") %>%
select(-vacc_dose2_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)

# write everything
t_desc_dose2b_rate %>%
write_csv("request-out/t_desc_dose2b_rate.csv")
```

### Table 5.2: Descriptive summaries for dose 2 AZ

- Similar to Table 5.1a, just subsetted to be those who recieved AZ for
  their second dose

```{r}
t_desc_dose2b_rate %>%
filter(vacc_dose2_name == "AZ") %>%
filter(xvar != "health_board") %>%
select(-vacc_dose2_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)
```

### Table 5.3: Descriptive summaries for dose 2 PB

- Similar to Table 5.1a, just subsetted to be those who recieved PB for
  their second dose

```{r}
t_desc_dose2b_rate %>%
filter(vacc_dose2_name == "PB") %>%
filter(xvar != "health_board") %>%
select(-vacc_dose2_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)
```

### Table 5.4: Poisson rate ratios

- Values are fully-adjusted estimated Poisson rate ratios, no suppression needed
- Wald 95% confidence intervals
- Reference levels are:
    + `bmi_cat = 18.5-24.9`
    + `vacc_dose_interval = dose2_day014`
    + `dose2_prior_infection_cat = No prior infection`
    + `wimd2019_quintile = 5least`
- The following CSVs are the raw model output version of this table:
    + `t_pois_coef_dose2.csv`
    + `t_pois_coef_dose2_az.csv`
    + `t_pois_coef_dose2_pb.csv`

```{r}
t_coef_dose2 <-
    qread("results/t_coef_dose2.qs") %>%
    tidy_coef() %>%
    rename(dose2 = coef)

t_coef_dose2_az <-
    qread("results/t_coef_dose2_az.qs") %>%
    tidy_coef() %>%
    rename(dose2_az = coef)

t_coef_dose2_pb <-
    qread("results/t_coef_dose2_pb.qs") %>%
    tidy_coef() %>%
    rename(dose2_pb = coef)

# print coef
t_coef_dose2 %>%
left_join(t_coef_dose2_az, by = c("xvar", "xlbl")) %>%
left_join(t_coef_dose2_pb, by = c("xvar", "xlbl")) %>%
filter(xvar != "bg_interval") %>%
filter(xvar != "health_board") %>%
kable() %>%
kable_styling(bootstrap_options = "striped", full_width = FALSE)

# save raw coef
tidy_raw_coef <- . %>%
    mutate(estimate = log(estimate)) %>%
    select(-conf_low, -conf_high)

qread("results/t_coef_dose2.qs") %>%
    tidy_raw_coef() %>%
    write_csv("request-out/t_pois_coef_dose2.csv")

qread("results/t_coef_dose2_az.qs") %>%
    tidy_raw_coef() %>%
    write_csv("request-out/t_pois_coef_dose2_az.csv")

qread("results/t_coef_dose2_pb.qs") %>%
    tidy_raw_coef() %>%
    write_csv("request-out/t_pois_coef_dose2_pb.csv")

```


# 6. Analysis of booster break through from 20 Dec 2021 (Omicron)

### Table 6.1: Descriptive summaries for Total Dose 2

- N, person-years, event counts, and rates are omitted if N or event count is between 1 and 9
- N and event counts rounded to nearest 10
- Person-years and rate per 1,000 person years rounded to 1dp
- The following CSV is the raw version of this table: `t_desc_doseb_omicron_rate.csv`

```{r}
t_desc_doseb_omicron_rate <-
    qread("results/t_desc_doseb_omicron_rate.qs") %>%
    mutate(
        xvar = fct_inorder(xvar),
        xlbl = fct_inorder(xlbl)
    )

# add totals to table
t_desc_doseb_omicron_total <-
    t_desc_doseb_omicron_rate %>%
    group_by(xvar, xlbl) %>%
    summarise(
        n = sum(n),
        pyears = sum(pyears),
        event = sum(event)
    ) %>%
    ungroup() %>%
    mutate(.after = xlbl, vacc_doseb_name = "Total") %>%
    mutate(rate = event / pyears * 1000)

t_desc_doseb_omicron_rate <-
    bind_rows(
        t_desc_doseb_omicron_total,
        t_desc_doseb_omicron_rate,
    )

# suppression
t_desc_doseb_omicron_rate <-
    t_desc_doseb_omicron_rate %>%
    mutate(
        n      = ifelse(1 <= n & n <= 9, NA, n),
        n      = round(n, -1),
        event  = ifelse(1 <= event & event <= 9, NA, event),
        event  = round(event, -1),
        pyears = ifelse(is.na(n), NA, pyears),
        pyears = round(pyears, 1),
        rate   = ifelse(is.na(event) | is.na(pyears), NA, rate),
        rate   = round(rate, 1)
    )

# print Total
t_desc_doseb_omicron_rate %>%
filter(vacc_doseb_name == "Total") %>%
filter(xvar != "health_board") %>%
select(-vacc_doseb_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)

# write everything
t_desc_doseb_omicron_rate %>%
write_csv("request-out/t_desc_doseb_omicron_rate.csv")
```

### Table 6.2: Descriptive summaries for MD booster bose

- Similar to Table 6.1, just subsetted to be those who recieved MD for
  their booster dose

```{r}
t_desc_doseb_omicron_rate %>%
filter(vacc_doseb_name == "MD") %>%
filter(xvar != "health_board") %>%
select(-vacc_doseb_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)
```

### Table 6.3: Descriptive summaries for PB booster dose

- Similar to Table 6.1, just subsetted to be those who recieved PB for
  their booster dose

```{r}
t_desc_doseb_omicron_rate %>%
filter(vacc_doseb_name == "PB") %>%
filter(xvar != "health_board") %>%
select(-vacc_doseb_name) %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)
```


### Table 6.4: Poisson rate ratios

- Values are fully-adjusted estimated Poisson rate ratios, no suppression needed
- Wald 95% confidence intervals
- Reference levels are:
    + `bmi_cat = 18.5-24.9`
    + `vacc_dose_interval = doseb_day014`
    + `dose2_prior_infection_cat = No prior infection`
    + `wimd2019_quintile = 5least`
- The following CSVs are the raw model output version of this table:
    + `t_pois_coef_doseb.csv`
    + `t_pois_coef_doseb_md.csv`
    + `t_pois_coef_doseb_pb.csv`

```{r}
t_coef_doseb <-
    qread("results/t_coef_doseb.qs") %>%
    tidy_coef() %>%
    rename(doseb = coef)

t_coef_doseb_md <-
    qread("results/t_coef_doseb_md.qs") %>%
    tidy_coef() %>%
    rename(doseb_md = coef)

t_coef_doseb_pb <-
    qread("results/t_coef_doseb_pb.qs") %>%
    tidy_coef() %>%
    rename(doseb_pb = coef)

# print coef table
t_coef_doseb %>%
left_join(t_coef_doseb_md, by = c("xvar", "xlbl")) %>%
left_join(t_coef_doseb_pb, by = c("xvar", "xlbl")) %>%
filter(xvar != "bg_interval") %>%
filter(xvar != "health_board") %>%
kable() %>%
kable_styling(bootstrap_options = "striped", full_width = FALSE)

# save raw coef
tidy_raw_coef <- . %>%
    mutate(estimate = log(estimate)) %>%
    select(-conf_low, -conf_high)

qread("results/t_coef_doseb.qs") %>%
tidy_raw_coef() %>%
write_csv("request-out/t_pois_coef_doseb.csv")

qread("results/t_coef_doseb_md.qs") %>%
tidy_raw_coef() %>%
write_csv("request-out/t_pois_coef_doseb_md.csv")

qread("results/t_coef_doseb_pb.qs") %>%
tidy_raw_coef() %>%
write_csv("request-out/t_pois_coef_doseb_pb.csv")
```


# 7. Analysis of booster break through for QCovid items from 20 Dec 2021 (Omicron)

### Table 7.1: Descriptive summaries

- N, person-years, event counts, and rates are omitted if N or event count is between 1 and 9
- If either the no or yes response are omitted (xlbl: 0 or 1) for a QCovid item (xvar), then both are omitted
- N and event counts rounded to nearest 10
- Person-years and rate per 1,000 person years rounded to 1dp
- The following CSV is the raw version of this table: `t_desc_doseb_qcovid_rate.csv`

```{r}
t_desc_doseb_qcovid_rate <-
    qread("results/t_desc_doseb_qcovid_rate.qs") %>%
    # suppress a row
    mutate(
        n      = ifelse(1 <= n & n <= 9, NA, n),
        n      = round(n, -1),
        event  = ifelse(1 <= event & event <= 9, NA, event),
        event  = round(event, -1),
        pyears = ifelse(is.na(n), NA, pyears),
        pyears = round(pyears, 1),
        rate   = ifelse(is.na(event) | is.na(pyears), NA, rate),
        rate   = round(rate, 1)
    ) %>%
    # suppress an xvar if one row has been suppressed
    group_by(xvar) %>%
    mutate(
        mis_n      = any(is.na(n)),
        mis_pyears = any(is.na(pyears)),
        mis_event  = any(is.na(event)),
        mis_rate   = any(is.na(rate))
    ) %>%
    ungroup() %>%
    mutate(
        n      = ifelse(mis_n,      NA, n),
        pyears = ifelse(mis_pyears, NA, pyears),
        event  = ifelse(mis_event,  NA, event),
        rate   = ifelse(mis_rate,   NA, rate)
    ) %>%
    select(-matches("mis_"))

t_desc_doseb_qcovid_rate %>%
kable(
    digits = c(0,0, 0,1,0,1),
    format.args = list(big.mark = ",")
) %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)

t_desc_doseb_qcovid_rate %>%
write_csv("request-out/t_desc_doseb_qcovid_rate.csv")
```

### Table 7.2: Poisson rate ratios

- Numbers are demographic-adjusted estimated rate ratios, no suppression needed
- Adjusted for:
    + vacc_dose_interval
    + sex
    + age_5y_cat
    + wimd2019_quintile
    + urban_rural_class
    + health_board
    + test_pre_dose2_cat
    + dose2_prior_infection_cat
    + vacc_dose1_dose2_diff_cat
    + bmi_cat
    + health_board
    + bg_interval
- The csv `t_pois_coef_doseb_qcovid.csv` is the raw model output that relates
  to this table


```{r}
# print coef table
qread("results/t_doseb_qcovid_rr.qs") %>%
rename(xvar = term) %>%
mutate(xlbl = 1) %>%
arrange(desc(estimate)) %>%
tidy_coef() %>%
select(-xlbl) %>%
kable() %>%
kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE
)

# save raw coef
qread("results/t_doseb_qcovid_rr.qs") %>%
mutate(estimate = log(estimate)) %>%
select(-conf_low, -conf_high) %>%
write_csv("request-out/t_pois_coef_doseb_qcovid.csv")
```


```{r}
beep()
```
